---
title: "Foundations in R pt 2 Assignment"
author: "Nikitha Mannem"
date: "2026-10-01"
output: html_document
---

# Set up
1. Load the tidyverse library. 
```{r}
library(tidyverse)
```

2. Load the billboard data. 
```{r}
data("billboard")
```

3. Describe the dataset. 
```{r}
summary(billboard)
```


4. Print the column names from this dataset. 
```{r}
colnames(billboard)
```


5. Print the first 5 entries in the dataset. 
```{r}
head(billboard, 5)
```


# Practice
1. Create a subset of data that has Jay-Z as the only artist. How many songs are in it?
```{r}
subset(billboard, artist == "Jay-Z") -> Jay_Z_data
Jay_Z_data$track
```

2. In the original dataset, use filter to find information about the song "All the small things." What was the highest position (i.e. lowest number) this song reached on the billboard 100? 
**the highest position this song reached was number 6 in week 12**
```{r}
subset(billboard, track == "All The Small Things")
```

3. Create a subset of data for songs Tim McGraw sang. Summarize the data for wk 1.
```{r}
subset(billboard, artist == "McGraw, Tim") -> TMG_data
TMG_data$track -> TMG_songs
TMG_songs
summary(TMG_data$wk1)
```

4. Create a function that can be used to determine if Britney Spears sang the song. 
```{r}
Is_it_Britney <- function(artist){billboard$artist == "Spears, Britney"}
Is_it_Britney(billboard$track)[5]
check_track <- billboard$track
Is_it_Britney(check_track)[272]
```

5. Test your function on the billboard data. Save the results to a new column called "wasItBritney"
```{r}
billboard$wasItBritney <- NA
Is_it_Britney(billboard$track) ->billboard$wasItBritney

billboard$wasItBritney
```

6. Use a for loop to determine if a song placed 50 or below or placed above 50 during the first week. 
```{r}
billboard$placement <- NA

for(i in 1:nrow(billboard)){if(billboard$wk1[i] >= 50){billboard$placement[i] <- "below"}
  else{billboard$placement[i] <- "above"}
}
```

7. Use an if else statement to determine the same information as you obtained with the for loop.
```{r}
billboard$placement2 <- NA

billboard$placement2 <- ifelse(billboard$wk1>=50, "below", "above")
```

8. Compare your results in 6 and 7. 
**they show the same values**
```{r}
print(billboard$placement)
print(billboard$placement2)
```

